---
title: "Statistical Analysis, Saliva Markers GRINTA!"
author: "Marc A.T. Teunis"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  error = FALSE,
  message = FALSE
                      
  )
```

# Session info
for reproducibility reasons
```{r, session}
sessioninfo::session_info()
```

# Introduction

The experimental statistical design here is a matched pairs experiment in which each subject recieves the treatment (variable `protocol` P1 - P5 for GRINTA! and protocol P1, P2, P4, P6 for TEMPO!). Subjects were randomly assigned to the order in which the protocols were executed. Each subject (in GRINTA! recieved P1 as first treatment: partially randomized block design). In TEMPO! the design was a fully randomized block. During each protocol subjects were sampled over a number of consistent time points (`time` variable) 

To account for this design we analyze the data first with a linear model (`lm`), then with a generalized linear model (`gls()`) and with a linear mixed effects model (`lme()`). All models are compared. AIC, visual inspection of distributions of the data and model residuals, QQ plots was used to select a transformation and the best fitting model. In the analysis below we focus on the 'GRINTA!' study exclusively.  

# Data preparation and exploratory data analysis

## Packages
```{r, packages}
library(tidyverse)
library(gramlyr)
library(pacman)
library(arm)
library(stargazer)
library(gdata)
library(RCurl)
library(readxl)
library(RColorBrewer)
library(wesanderson)
# install.packages("docxtools")
library(docxtools)
library(broom)
library(nlme)
#library(lme4)
library(salivar)
```

## Data 

### Load tidy dataset
There are two datasets in this package, one containing the raw data of the study (measured in saliva, and for comparison, a few serum markers), and one containing meta data for the measured analytes.

```{r, data_load}
# data(package = "salivar")
data(package = "salivar", dataset = "data_all_tidy")
data(package = "salivar", dataset = "analyte_annotations")
data(package = "salivar", dataset = "data_order_tidy")

data_all_tidy
data_order_tidy
sum(is.na(data_all_tidy))
## check factor levels
map(data_all_tidy[, c(1:4)], unique)
map(data_order_tidy, levels)
```

### Remove subject 17 and 15, which is not valid?
**@Shirley: Do you know if the data for volunteer 17 should be removed, or that the order for this volunteer is missing from the 'order' dataframe**
```{r, remove_subject_17, eval=FALSE}
data_all_tidy <- data_all_tidy %>%
  dplyr::filter(subject != "17", ) %>% drop.levels()

```

### Join analyte data with order data
```{r, join_data}
data_all_tidy <- left_join(data_all_tidy, data_order_tidy) %>%
  droplevels() %>%
  mutate(
    protocol = as_factor(protocol),
    order = as_factor(order),
    subject = as_factor(subject)
  )

map(data_all_tidy[, -5], unique)
```

### Nest by analyte
```{r, nest}
## log transform concentration and nest data for analyte
## chack if there are zeros
data_all_tidy %>%
  ggplot(aes(x = concentration+1)) + ## ading 1 to all as pseudo
  geom_histogram()

data_nested <- data_all_tidy %>%
  mutate(concentration_log = log(concentration+1)) %>% ## adding 1 to all as pseudo
  group_by(analyte) %>%
  nest()

## set names $data
names(data_nested$data) <- data_nested$analyte

```

```{r, implicit_missingness, eval=FALSE, include=FALSE}
### Complete data for implicit missing values (combinations that are not present in the data)
subjects <- length(unique(data_slpi$subject))
protocols <- length(unique(data_slpi$protocol))   
times <- length(unique(data_slpi$time))   

n_rows_expect <- subjects * protocols * times

```

### Function to calculate expected rows per analyte, and fill if neccessary
```{r, fun_missingness}
expect_rows <- function(df){
  
  subjects <- length(unique(df$subject))
  protocols <- length(unique(df$protocol))   
  times <- length(unique(df$time))   

  n_rows_expect <- subjects * protocols * times
  return(n_rows_expect)
}

data_nested <- data_nested %>%
  dplyr::mutate(
    actual_rows = map_int(data, nrow),
    expected_rows = map_int(data, expect_rows))

all <- data_nested$actual_rows == data_nested$expected_rows

stopifnot(
all(all)
)

```

**Conclusion:** No implicity missing values are present in the data so no further steps to complement the missingness are neccessary.

### Missingness, visual
```{r, visual_mssingness}
data_nested <- data_nested %>%
  mutate(visual_missingness = map(
    data, naniar::vis_miss)
    )
names(data_nested$visual_missingness) <- data_nested$analyte
data_nested$visual_missingness

## add analytes back to data frames
data_nested <- data_nested %>%
  dplyr::mutate(
    data = map2(
      .x = data,
      .y = analyte,
      .f = salivar::add_grouping_value_back_to_list_column_df
    )
  )

## check
# data_nested$data[1:2]
```

## Distributions
```{r, distributions}

## add to nested dataframe, before and after log transformation
data_nested <- data_nested %>%
  mutate(histograms_conc = map2(
    .x = data,
    .y = analyte,
    .f = plot_histogram_from_nested_df,
    var = "concentration"
  ),
        histograms_conc_log = map2(
    .x = data,
    .y = analyte,
    .f = plot_histogram_from_nested_df,
    var = "concentration_log"
  ))

names(data_nested$histograms_conc) <- data_nested$analyte
names(data_nested$histograms_conc_log) <- data_nested$analyte
```

### Before log transform
```{r, before_log}
data_nested$histograms_conc$SLPI
data_nested$histograms_conc$mmp9.se
```

### After log transform
```{r, after_log}
data_nested$histograms_conc_log$SLPI
data_nested$histograms_conc_log$mmp9.se
```

Conclusion: most of the distributions appear to approach a normal distribution, after log transformation, some do display a deviance from a normal distribution.

### Outliers, per subject
```{r}
data_nested <- data_nested %>%
  mutate(boxplots_per_subject = map2(
    .x = data,
    .y = analyte,
    .f = plot_boxplot_from_nested_df,
    var_x = "subject",
    var_y = "concentration_log"
  ))
i = 1:nrow(data_nested)
data_nested$boxplots_per_subject[i]

```

## QQ-Norm and Shapiro tests
**TODO**
```{r}

```

# Statistical models

## Resources

https://www.youtube.com/watch?v=nPdrWq_Sb-U
by Erin Buchanan

http://rstudio-pubs-static.s3.amazonaws.com/25354_2eeb42bf56d44cbc9ee217535aa962aa.html

https://stats.stackexchange.com/questions/13784/repeated-measures-anova-with-lme-lmer-in-r-for-two-within-subject-factors

https://stats.stackexchange.com/questions/38995/using-lme-to-analyse-a-complete-randomized-block-design-with-repeated-measures/39180#39180

## Function to apply different models to the data
This function is a core function of the `{salivar}` package and a wrapper to multiple modelling function (`lm()`, `gls()` and `lme()`). 
```{r}
?salivar::model_me
```

## List of model definitions
```{r}
## lm models
models_lm <- list(
  model_lm_base = concentration_log ~ 1,
  model_lm_protocol = concentration_log ~ protocol,
  model_lm_protocol_time = concentration_log ~ protocol * time
)

## gls models
models_gls <- list(
  # intercept only
  model_gls_base = concentration_log ~ 1, 
  model_gls_protocol = concentration_log ~ protocol,
  model_gls_protocol_time = concentration_log ~ protocol * time
  )    

## mixed-effects models
models_lme <- list(
  # intercept only
  model_lme_base = concentration_log ~ 1,
  model_lme_protocol = concentration_log ~ protocol,  
  model_lme_protocol_time = concentration_log ~ protocol + time,
  model_lme_protocol_time_int = concentration_log ~ protocol * time 

  )    

```


```{r, include=FALSE, eval=FALSE}
## Test models (computation time & optimization)
## test data (one analyte which we know shows significant changes over time and over protocols)
model_me_safe <- safely(salivar::model_me)

model_1 <- model_me_safe(
  df = data_slpi_saliva,
  formula = models_lme$model_lme_protocol_time_int,
  type = "lme",
  random = ~1 | subject,
      method = "ML", 
      na.action = "na.omit"
  
)

model_2 <- model_me_safe(
  df = data_slpi_saliva,
  formula = models_lme$model_lme_protocol_time_int,
  type = "lme",
  random = ~1 | order/subject,
      method = "ML", 
      na.action = "na.omit"
  
)

model_3 <- model_me_safe(
  df = data_slpi_saliva,
  formula = models_lme$model_lme_protocol_time_int,
  type = "lme",
  random = ~1 | subject/protocol,
      method = "ML", 
      na.action = "na.omit"
  
)

model_1$result %>% summary %>% pluck("AIC")
model_2$result %>% summary %>% pluck("AIC")
model_3$result %>% summary %>% pluck("AIC")

model_1$result %>% resid() %>% plot
model_2$result %>% resid() %>% plot
model_3$result %>% resid() %>% plot

model_1$

anova(model_1$result,
      model_2$result
  
)

```

## Apply models to data
```{r}
data_nested %>% 
  mutate(
## lm models
      model_lm_base = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lm$model_lm_base,
      type = "lm"),
    summary_lm_base = map(
      .x = model_lm_base,
      .f = summary),
    model_lm_protocol = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lm$model_lm_protocol,
      type = "lm"),
    summary_lm_protocol = map(
      .x = model_lm_protocol,
      .f = summary),
    model_lm_protocol_time = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lm$model_lm_protocol_time,
      type = "lm"),
    summary_lm_protocol_time = map(
      .x = model_lm_protocol_time,
      .f = summary),
## gls models
    model_gls_base = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_gls$model_gls_base,
      type = "lm"),
    summary_gls_base = map(
      .x = model_gls_base,
      .f = summary),
## mixed-effects models (lme)
    model_lme_base = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lme$model_lme_base,
      type = "lme",
      random = ~1 | subject,
      method = "ML", 
      na.action = "na.omit"
      ),
    summary_lme_base = map(
      .x = model_lme_base, 
      .f = summary
    ),
    model_lme_protocol = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lme$model_lme_protocol,
      type = "lme",
      random = ~1 | subject,
      method = "ML", 
      na.action = "na.omit"
      ),
    summary_lme_protocol = map(
      .x = model_lme_protocol, 
      .f = summary
    ),
    model_lme_protocol_time = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lme$model_lme_protocol_time,
      type = "lme",
      random = ~1 | subject,
      method = "ML", 
      na.action = "na.omit"
      ),
    summary_lme_protocol_time = map(
      .x = model_lme_protocol_time, 
      .f = summary
      ),
    model_lme_protocol_time_int = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lme$model_lme_protocol_time_int,
      type = "lme",
      random = ~1 | subject,
      method = "ML", 
      na.action = "na.omit"
      ),
    summary_lme_protocol_time_int = map(
      .x = model_lme_protocol_time_int, 
      .f = summary
    ),
    model_lme_protocol_time_full = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lme$model_lme_protocol_time_int,
      type = "lme",
      random = ~1 | protocol/subject,
      method = "ML", 
      na.action = "na.omit"
      ),
    summary_lme_protocol_time_full = map(
      .x = model_lme_protocol_time_full, 
      .f = summary
    ),
    model_lme_protocol_time_full_order = map(
      .x = data,
      .f = salivar::model_me,
      formula = models_lme$model_lme_protocol_time_int,
      type = "lme",
      random = ~1 | order/subject,
      method = "ML", 
      na.action = "na.omit"
      ),
    summary_lme_protocol_time_full_order = map(
      .x = model_lme_protocol_time_full_order, 
      .f = summary
    )
) -> data_nested
```

## Get model AIC for lme models only
```{r}

data_nested <- data_nested %>%
  mutate(aic_lme_base = map_dbl(summary_lme_base, pluck, "AIC"),
         aic_lme_protocol = map_dbl(summary_lme_protocol, pluck, "AIC"),
         aic_lme_protocol_time = map_dbl(summary_lme_protocol_time, pluck, "AIC"),
         aic_lme_protocol_time_int = map_dbl(summary_lme_protocol_time_int, pluck, "AIC"),
         aic_lme_full = map_dbl(summary_lme_protocol_time_full, pluck, "AIC"),
         aic_lme_full_order = map_dbl(summary_lme_protocol_time_full_order, pluck, "AIC")
         )

data_aic <- data_nested %>%
  dplyr::select(analyte, aic_lme_base:aic_lme_full_order) %>%
  pivot_longer(cols = c(aic_lme_base:aic_lme_full_order), 
               names_to = "model", 
               values_to = "aic") 

data_lme_full_order_aic <- data_aic %>%
  dplyr::filter(model == "aic_lme_full_order")

data_aic %>%
  ggplot(aes(x = reorder(as_factor(model), aic), y = aic)) +
  geom_point() +
  geom_point(data = data_lme_full_order_aic, 
             aes(x = reorder(as_factor(model), aic), y = aic), colour = "red") +
  coord_flip() +
  facet_wrap(~analyte)

```

## What is the best model to fit the data
From the graph above we can select the model with lowest AIC, the model with order and subject as random intercepts and protocol, time and their interaction as fixed factors seems to have the lowest AIC for most analytes. The AIC information criterium for this model is indicated in red in the above graph.

```{r, eval=FALSE, include=FALSE}
### Compare models with anova
model_comparisons_lme <- list()

for(i in seq_along(1:nrow(data_nested))) {

  model_comparisons_lme[[i]] <-
salivar::compare_models(model_list = list(
  data_nested$model_lme_base[[i]],
  data_nested$model_lme_protocol[[i]],
  data_nested$model_lme_protocol_time[[i]],
  data_nested$model_lme_protocol_time_int[[i]],
  data_nested$model_lme_protocol_time_full[[i]]
))

}

## add model comparisons to nested df
data_nested$lme_comparisons <- model_comparisons_lme

plot_residuals_from_model(model = data_nested$model_lme_protocol_time_full[[26]])
plot_residuals_from_model(model = data_nested$model_lme_protocol_time_int[[26]])

data_nested$model_lme_protocol_time_full[[26]] %>% resid() %>% shapiro.test()
data_nested$model_lme_protocol_time_full[[26]] %>% resid() %>% hist()
  
```

## Select only the best model (with overall lowest AIC) from the data for further calculations.

```{r}
best_model <- "model_lme_protocol_time_full_order"

names(data_nested)
data_nested_best_model <- data_nested %>%
  dplyr::select(
    analyte,
    data,
    model_lme_protocol_time_full_order,
    summary_lme_protocol_time_full_order
    )


```

# The results for the best model
```{r}
data_nested_best_model <- data_nested_best_model %>% 
  dplyr::mutate(results_models = map(model_lme_protocol_time_full_order, 
                                     salivar::extract_results_from_model,
                                     scientific = FALSE,
                                     digits = 4),
                results_models = map2(
                  .x = results_models,
                  .y = analyte,
                  .f = salivar::add_grouping_value_back_to_list_column_df
                ))


name_values <- data_nested_best_model$analyte
data_nested_best_model[, c(2:ncol(data_nested_best_model))] <- map(
  .x = data_nested_best_model[, c(2:ncol(data_nested_best_model))],
  .f = salivar::set_names_on_list_column,
  name_values = name_values
)

```

## The three analytes with most significant model parameters
```{r}
data_nested_best_model$results_models$mmp9.se
data_nested_best_model$results_models$SLPI
data_nested_best_model$results_models$`MMP-9`
```

## Plotting the significant results of the models for each analyte
```{r}
# TODO; create an S3 class (salivar_model_result) for the object that is input to this function 
salivar_model_result = data_nested_best_model$results_models[[1]]

## all result plots
names(data_nested_best_model)
data_nested_best_model <- data_nested_best_model %>%
  mutate(result_plot = map(
    .x = results_models,
    .f = plot_model_results,
    significance = 0.05,
    colour = "red"
  ))

data_nested_best_model$result_plot
```

```{r, eval=FALSE}
## save to disk
paths <- map(
  .x = paste(
         paste0("inst/images/", str_replace_all(data_nested_best_model$analyte,
                pattern = "\\.", replacement = "_")), 
       "significant_p.png", sep = "_"),
  .f = file.path()
)

paths  

walk2(
  .x = paths,
  .y = data_nested_best_model$result_plot,
  .f = ggsave,
  width = 8, height = 12, units = "cm", dpi = 300
)

```

## Residual plots
```{r}

model = data_nested_best_model$model_lme_protocol_time_full_order$mmp9.se
analyte = "mmp9.se"

plot_fitted_vs_residuals_scatter <- function(model, analyte, ...){
  

  resids <- stats::residuals(model) %>% 
    tibble::as_tibble()
  fitted <- stats::fitted(model) %>% 
    tibble::as_tibble()

  df <- dplyr::bind_cols(resids,
                         fitted)
  
  
 plot_scatter <- df %>% 
   ggplot(aes(x = value1,
              y = value)) +
   geom_point(shape = 1) +
   ggtitle(paste(analyte, "\n", model$call)) +
   geom_hline(yintercept = 0, ...) +
   xlab("Fitted") +
   ylab("Residuals")
 
 return(plot_scatter) 
} 

## distribution of residuals
plot_fitted_vs_residuals_histogram <- function(model, analyte, ...){
  
  resids <- stats::residuals(model) %>% 
    tibble::as_tibble()
  fitted <- stats::fitted(model) %>% 
    tibble::as_tibble()

  df <- dplyr::bind_cols(resids,
                         fitted)
  
  shap <- shapiro.test(df$value)
  shap_result <- paste(shap$method, "p-value:", round(shap$p.value, digits = 4))
  
  plot_histogram <- df %>% 
    ggplot(aes(x = value)) +
    geom_histogram() +
    ggtitle(paste(analyte, "\n", model$call)) +
    annotate("text", label= shap_result, y = 10, x = 0, colour = "red")
    
#   geom_hline(yintercept = 0, ...) +
  
 return(plot_histogram) 
} 


## tests
plot_fitted_vs_residuals_scatter(model = model,
                         analyte = analyte,
                         colour = "red", linetype = "dashed",
                         size = 1)

plot_fitted_vs_residuals_histogram(model = model,
                         analyte = analyte,
                         colour = "red", linetype = "dashed",
                         size = 1)


## add residual plots to nested dataframe
data_nested_best_model <- data_nested_best_model %>%
  mutate(residual_plots = map2(
    .x = model_lme_protocol_time_full_order,
    .y = analyte,
    .f = plot_fitted_vs_residuals_scatter
  ))

data_nested_best_model <- data_nested_best_model %>%
  mutate(residual_histogram_plots = map2(
    .x = model_lme_protocol_time_full_order,
    .y = analyte,
    .f = plot_fitted_vs_residuals_histogram
  ))


data_nested_best_model$residual_plots
data_nested_best_model$residual_histogram_plots
```

```{r, eval=FALSE}

## save residual scatter plots to disk
paths <- map(
  .x = paste(
         paste0("inst/images/", str_replace_all(data_nested_best_model$analyte,
                pattern = "\\.", replacement = "_")), 
       "residuals_best_model.png", sep = "_"),
  .f = file.path()
)

paths  


walk2(
  .x = paths,
  .y = data_nested_best_model$residual_plots,
  .f = ggsave,
  width = 12, height = 8, units = "cm", dpi = 300
)

```

```{r, eval=FALSE}
## save histograms to disk
paths <- map(
  .x = paste(
         paste0("inst/images/", str_replace_all(data_nested_best_model$analyte,
                pattern = "\\.", replacement = "_")), 
       "residuals_histograms_best_model.png", sep = "_"),
  .f = file.path()
)

paths  

walk2(
  .x = paths,
  .y = data_nested_best_model$residual_histogram_plots,
  .f = ggsave,
  width = 12, height = 8, units = "cm", dpi = 300
)

## Create analyte panel

```


```{r, include=FALSE, eval=FALSE}
## Post hoc analysis 
library(emmeans)

names(data_nested_best_model$summary_lme_protocol_time_full_order) <- 
  data_nested_best_model$analyte
data_nested_best_model$summary_lme_protocol_time_full_order$mmp9.se
data_nested_best_model$summary_lme_protocol_time_full_order[[24]]

## test on one model
model_test <- data_nested_best_model$model_lme_protocol_time_full_order[[24]]
lsd <- emmeans(model_test, pairwise ~ protocol:time)
summary(lsd, adjust = "Tukey") 
number_of_analytes <- 

statistics <- statistics %>%
  dplyr::mutate(`p-value_adjust` = p.adjust(`p-value`, n = 27))

statistics$`p-value_adjust`[1:5]

results <- statistics %>%
  dplyr::select(group, param, `p-value`, `p-value_adjust`)
results
```

## Multilevel design

- Factors: Protocol -> Time
- Predictors: Protocol / Time 
- Dependent variable: concentration
- random effects: order / subject

The design is partially randomized cross-over experiment, but does the current model reflect this? I need to consult with a statistician to learn if the current model has the most optimal settings to fit the data. According the AIC, and BIC criteria  the current model used: 

```
lme(concentration_log ~ time * protocol, 
  error = +1 | order/subject )
```

This is a random intercept model with fixed predictors: protocol (type of exercise) and time (sample collection time) and random effects factors order (order in which the subject cycled a protocol) and subject (subject id) 

## Conclusions,

The significant model-parameters over sample time and protocol are tallied in the following table 
```{r}

data_nested_best_model$results_models$slpi.se

signi_tally <- data_nested_best_model$results_models %>%
  dplyr::bind_rows() %>%
  group_by(analyte) %>%
  dplyr::select(`p-value`, analyte) %>%
  dplyr::filter(`p-value` < 0.01) %>%
  tally %>%
  arrange(desc(n)) 

signi_tally %>%
  knitr::kable()

```

# Image panels per analyte

 - One panel with all visualizations per analyte, for overview
 - Include models and model checks (residuals, p-values, histograms, qqplot)

# PCA analysis
Generate a normalized version of the data (normalize t=0 to 100%) and express every other value as a percentage of t=0

## Normalize data
Did it work?

```{r}
data_nested <- data_nested %>%
  mutate(data_normalized_summary = map(
    .x = data,
    .f = create_normalized_summary   
        ),
        data_normalized_summary = map2(
          .x = data_normalized_summary,
          .y = analyte,
          .f = salivar::add_grouping_value_back_to_list_column_df
        ))


data_nested$data_normalized_summary[[1]]



## Check one line
# RColorBrewer::display.brewer.all()
palette <- RColorBrewer::brewer.pal(7, "Dark2")
palette_graph <- c("#000000", palette[c(1:3,4)])
image_directory <- here::here(
  "inst",
  "images",
  "normalized"
)

names(data_nested$data_normalized_summary) <- data_nested$analyte
```

## Normalized figures for: 

The data has been normalized according the following formula

$Yi.analyte.norm = \frac{mean(Y.analyte.t=0)}{Yi.analyte}$


 - MMP-serum
 - MMP-saliva
 - SLPI-serum
 - SLPI-saliva
 
```{r, fig.height=15, fig.width= 20}
## all normalized plots
data_nested <- data_nested %>%
  dplyr::mutate(normalized_plots = map2(
    .x = data_normalized_summary,
    .y = analyte,
    .f = salivar::draw_lines,
    palette_graph = palette_graph,
    f.width = 30, 
    f.height = 28
    ))
 
data_nested$normalized_plots <- 
  salivar::set_names_on_list_column(
    data_nested$normalized_plots, 
    name_values = data_nested$analyte
)

## only mmp and slpi for serum and saliva
data_nested$normalized_plots$mmp9.se
data_nested$normalized_plots$`MMP-9`
data_nested$normalized_plots$slpi.se
data_nested$normalized_plots$SLPI

```

## Load data
```{r, eval=FALSE}
library(tidyverse)
library(salivar)

data_citrulliner <- citrulliner::citrulline_data %>%
  dplyr::select(
    subject,
    protocol,
    time,
    analyte,
    concentration
  )

data_salivar <- salivar::data_all_tidy %>%
  dplyr::select(
    subject,
    protocol,
    time,
    analyte,
    concentration
  )
```

## PCA analysis
For the saliva data
```{r}
df = data_salivar

times <- unique(salivar::data_all_tidy$time)

pca_plot_list <- map(
  .x = times,
  .f = run_pca,
  df = df,
  na_filter = 30
)

pca_plot_list
```

```{r, include=FALSE, eval=FALSE}
## save to disk
paths <- map(
  .x = paste(
         paste0("inst/images/", "tis",times), 
       "pca_plot.png", sep = "_"),
  .f = file.path()
)

walk2(
  .x = paths,
  .y = pca_plot_list,
  .f = ggsave,
  device = "png",
  width = 12, height = 8, units = "cm", dpi = 300
)

```

## <mark> PCA citrulline / immune parameters </mark>

<mark> WORK IN PROGRESS </mark>


```{r, eval=FALSE}
## TODO

data_citrulliner_wide <- data_citrulliner %>%
  salivar::normalize_for_protocol_P1() %>%
  dplyr::select(
    subject, 
    protocol, 
    time, 
    analyte, 
    concentration_normalized
  ) %>% 
  tidyr::pivot_wider(names_from = analyte, 
                     values_from = concentration_normalized) 

data_citrulliner_wide <- data_citrulliner %>%
  unite(col = "keval", 
  c(subject, protocol, time), remove = FALSE)

data_citrulliner_wide <- data_citrucitrulliner_wide %>%
data_citrullinerame()

rownames(data_citrulliner_wide) <- data_citrulliner_wide$keval

unique(data_citrulliner_wide$keval) %>% length()

# data_citrullinermissing values, the step  below removes them

ddata_citrullinersalivar <- data_citrulliner_wide[ , data_citrulliner6)] %>% na.omit()
naniar::vis_miss(data_numeric_salivar)

pca_wide <- prcomp(data_numeric_salivar, center = TRUE, scale. = TRUE)

summary(pca_wide)

data_salivar_pca <- as.data.frame(pca_wide$x)  

data_salivar_pca <- data_salivar_pca %>%
  mutate(keyval = rownames(.)) %>%
  separate(col = "keyval", 
           into = c("subject", "protocol", "time"), 
           remove = FALSE)

plot_pca_salivar <- data_salivar_pca %>%
  ggplot(aes(x = PC1, 
             y = PC2)) +
  geom_point(aes(colour = protocol, shape = time)) + 
  theme_bw() 

plot_pca_salivar

pca_plot2 %>% ggsave(filename = 
                       here::here(
                         "inst",
                         "images",
                         "pca_plot_salivar.svg"))

```
